Exact Asymptotics for Causal Mediation Analysis

### 2024-06-18 notes/William Analysis.Rmd
Update consolidated notes

library(lme4)     # For fitting GLMMs using glmer()
library(MASS)     # For simulating multivariate normals using mvrnorm()
library(merDeriv) # For computing the full information matrix from models fit using lme4
library(tictoc)

@@

# Number of replicates for each sample size
# We use fewer here because model fitting takes longer
num_reps = 200
```

```{r MC-bin-bin-mix, include=FALSE, cache=TRUE}
# , dependson=c("boilerplate")}

@@

  # Store output for current values of n and K
  all_a_hats[[j]] = lapply(some_output, function(x) x$a_hat)
  all_theta_hats[[j]] = lapply(some_output, function(x) x$theta_hat)
  all_b_hats[[j]] = lapply(some_output, function(x) x$b_hat)
  all_gamma_hats[[j]] = lapply(some_output, function(x) x$gamma_hat)
  all_M_covs[[j]] = lapply(some_output, function(x) x$M_cov)
  all_Y_covs[[j]] = lapply(some_output, function(x) x$Y_cov)
  all_med_hats[[j]] = lapply(some_output, function(x) x$med_hat)
  all_med_SEs[[j]] = lapply(some_output, function(x) x$med_SE)
}

@@

# Save output
# save(all_a_hats, all_theta_hats, all_b_hats, all_gamma_hats, all_M_covs, all_Y_covs, all_med_hats, all_med_SEs, runtime, file = "William-Analysis_cache/MC-bin-bin-mix.RData")
# load("William-Analysis_cache/MC-bin-bin-mix.RData")
```

We now process the output from our simulation and summarize the results in a table. We give the mean and median of the estimated standard errors, as well as their relative errors (in $\%$) as estimates of the empirical standard error. See Table \ref{tab:results_bin_bin_fix}.

@@

# Combine into a table
# results = data.frame(Empirical = emp_SEs,
#                      Mean = mean_delta_SEs, Median = median_delta_SEs,
#                      "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
#                      "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)
# results = data.frame(n = all_Ns, Empirical = emp_SEs,
#                      Mean = mean_delta_SEs, Median = median_delta_SEs,
#                      "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
#                      "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)
results = data.frame(n = N_K_combs$N, K = N_K_combs$K, Empirical = emp_SEs,
                     Mean = mean_delta_SEs, Median = median_delta_SEs,
                     "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
                     "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)

@@

```{r results="asis", echo=FALSE}
## Print table
kbl(results, format = "latex", booktabs = T, caption = "Summary of SE estimates for mediation effect under continuous response, continuous mediator, fixed-effects model \\label{tab:results_bin_bin_fix}")
```

I don't really understand the results I'm getting here. It should help if I can compare my estimates of the mediation effect with the true value.

```{r}
load("../data/Exact_Asymptotics/Bin-Bin-Mix_Results.RData")

true_med = Phi_of_xi(a, theta, b, gamma, x_pred, W_pred)

load("William-Analysis_cache/latex/unnamed-chunk-37_44bd8d1152289e689fbff9185d17042c.RData", ex <- new.env())
ls.str(ex)
```

Investigate normality of estimators
```{r}
library(ggplot2)
library(magrittr)

med_hats_to_plot = unlist(data_med_hats)
Ns_to_plot = rep(N_K_combs$N, times = sapply(data_med_hats, length))
Ks_to_plot = rep(N_K_combs$K, times = sapply(data_med_hats, length))
data_med_hat_to_plot = data.frame(med_hat = med_hats_to_plot, N = Ns_to_plot, K = Ks_to_plot)

data_med_hat_to_plot_OLD = data_med_hat_to_plot
data_med_hat_to_plot %<>% dplyr::filter(med_hat < 4)

ggplot(data_med_hat_to_plot, aes(x = med_hat)) +
  # geom_histogram(bins = 30) +
  geom_density() +   theme_minimal() +
  facet_grid(rows = vars(N), cols = vars(K), scales = "free")

# QQ plot for each sample size
ggplot(data_med_hat_to_plot, aes(sample = med_hat)) +
  stat_qq() +
  stat_qq_line() +
  facet_grid(rows = vars(N), cols = vars(K), scales = "free")
  theme_minimal()

# Test normality in each sample size
all_norm_p_vals = c()
for(i in 1:nrow(N_K_combs)){
  p_val = shapiro.test(data_med_hats[[i]])$p.value
  all_norm_p_vals = c(all_norm_p_vals, p_val)
}

results$Shapiro_p = all_norm_p_vals
results

@@

>>>>

library(lme4)     # For fitting GLMMs using glmer()
library(MASS)     # For simulating multivariate normals using mvrnorm()
library(merDeriv) # For computing the full information matrix from models fit using lme4
library(tictoc)
library(ggplot2)
library(magrittr)
library(dplyr)

@@

# Number of replicates for each sample size
# We use fewer here because model fitting takes longer
num_reps = 200

# Easier version of problem
all_Ns = c(200, 500)
all_Ks = c(5, 20)

sigma_a_0 = 0.5
sigma_a_1 = 0.5 * abs(a_1)
cor_a0_a1 = 0.5
cov_a0_a1 = sigma_a_0 * sigma_a_1 * cor_a0_a1
theta = c(sigma_a_0, sigma_a_1, cor_a0_a1)
Sigma_a = matrix(c(sigma_a_0^2, cov_a0_a1, cov_a0_a1, sigma_a_1^2), nrow = 2, ncol = 2)

sigma_b_0 = 0.5 * abs(b_0)
sigma_b_1 = 0.5 * abs(b_1)
cor_b0_b1 = 0.5
cov_b0_b1 = sigma_b_0 * sigma_b_1 * cor_b0_b1
gamma = c(sigma_b_0, sigma_b_1, cor_b0_b1)
Sigma_b = matrix(c(sigma_b_0^2, cov_b0_b1, cov_b0_b1, sigma_b_1^2), nrow = 2, ncol = 2)
```

This simulation takes a while. As such, I have saved the results and don't re-run things every time I compile this document. To re-simulate, un-comment the following code chunk.

```{r MC-bin-bin-mix, include=FALSE}
# , dependson=c("boilerplate")}

@@

  # Store output for current values of n and K

  all_a_hats[[j]] = lapply(some_output, function(x) x$a_hat)
  all_theta_hats[[j]] = lapply(some_output, function(x) x$theta_hat)
  all_b_hats[[j]] = lapply(some_output, function(x) x$b_hat)
  all_gamma_hats[[j]] = lapply(some_output, function(x) x$gamma_hat)
  all_M_covs[[j]] = lapply(some_output, function(x) x$M_cov)
  all_Y_covs[[j]] = lapply(some_output, function(x) x$Y_cov)
  all_med_hats[[j]] = lapply(some_output, function(x) x$med_hat)
  all_med_SEs[[j]] = lapply(some_output, function(x) x$med_SE)
}

@@

# Save output
# save(N_K_combs, all_a_hats, all_theta_hats, all_b_hats, all_gamma_hats, all_M_covs, all_Y_covs, all_med_hats, all_med_SEs, runtime, file = "William-Analysis_cache/MC-bin-bin-mix.RData")
# load("D:/OneDrive/OneDrive - Simon Fraser University (1sfu)/Research/Postdoc/Exact_Asymptotics/notes/William-Analysis_cache/MC-bin-bin-mix.RData", verbose = TRUE)
```

We now process the output from our simulation and summarize the results in a table. We give the mean and median of the estimated standard errors, as well as their relative errors (in $\%$) as estimates of the empirical standard error. See Table \ref{tab:results_bin_bin_mix}.

@@

# Combine into a table
results = data.frame(n = N_K_combs$N, K = N_K_combs$K, Empirical = emp_SEs,
                     Mean = mean_delta_SEs, Median = median_delta_SEs,
                     "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
                     "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)

# Add extra values of n and K to list
# all_N_combs = c(N_K_combs$N)
# all_K_combs = c(N_K_combs$K)
# results = data.frame(n = all_N_combs, K = all_K_combs, Empirical = emp_SEs,
#                      Mean = mean_delta_SEs, Median = median_delta_SEs,
#                      "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
#                      "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)

@@

<!-- I'm now trying to figure out what's causing an error which only occurs on knit. Trying to isolate it in the following code chunk: -->

<!-- ```{r} -->
<!-- load("D:/OneDrive/OneDrive - Simon Fraser University (1sfu)/Research/Postdoc/Exact_Asymptotics/notes/William-Analysis_cache/MC-bin-bin-mix.RData", verbose = TRUE) -->

<!-- # print(all_med_hats) -->
<!-- ``` -->

```{r results="asis", echo=FALSE}
## Print table
kbl(results, format = "latex", booktabs = T, caption = "Summary of SE estimates for mediation effect under continuous response, continuous mediator, fixed-effects model \\label{tab:results_bin_bin_mix}")
```

<!-- I don't really understand the results I'm getting here. It should help if I can compare my estimates of the mediation effect with the true value. -->

<!-- ```{r} -->
<!-- load("../data/Exact_Asymptotics/Bin-Bin-Mix_Results.RData") -->

<!-- true_med = Phi_of_xi(a, theta, b, gamma, x_pred, W_pred) -->

<!-- load("William-Analysis_cache/latex/unnamed-chunk-37_44bd8d1152289e689fbff9185d17042c.RData", ex <- new.env()) -->
<!-- ``` -->

There's lots that I don't understand about the above results. One think I decided to explore is the normality of the estimators. To this end, I make some density plots, and also perform Shapiro-Wilk tests for normality. See Table \ref{tab:results_bin_bin_mix-Shapiro} for the test results.

```{r}
med_hats_to_plot = unlist(data_med_hats)
Ns_to_plot = rep(N_K_combs$N, times = sapply(data_med_hats, length))
Ks_to_plot = rep(N_K_combs$K, times = sapply(data_med_hats, length))
data_med_hat_to_plot = data.frame(med_hat = med_hats_to_plot, N = Ns_to_plot, K = Ks_to_plot)

data_med_hat_to_plot_OLD = data_med_hat_to_plot
data_med_hat_to_plot %<>% dplyr::filter(med_hat < 4)
```

```{r}
ggplot(data_med_hat_to_plot, aes(x = med_hat)) +
  geom_density() +
  facet_grid(rows = vars(N), cols = vars(K), scales = "free") +
  ggtitle("Density of mediation effect estimates")
  # theme_minimal()
```

```{r}
# QQ plot for each sample size
ggplot(data_med_hat_to_plot, aes(sample = med_hat)) +
  stat_qq() +
  stat_qq_line() +
  facet_grid(rows = vars(N), cols = vars(K), scales = "free") +
  ggtitle("QQ plot of mediation effect estimates")
  # theme_minimal()
```

```{r}
# Test normality in each sample size
all_norm_p_vals = c()
for(i in 1:nrow(N_K_combs)){
  p_val = shapiro.test(data_med_hats[[i]])$p.value
  all_norm_p_vals = c(all_norm_p_vals, p_val)
}

results$Shapiro_p = all_norm_p_vals
# results

# # Check how p-value varies with N and K
# p_vals_by_N = results %>% group_by(n) %>% summarize(mean_p = mean(Shapiro_p), sd_p = sd(Shapiro_p))
# p_vals_by_K = results %>% group_by(K) %>% summarize(mean_p = mean(Shapiro_p), sd_p = sd(Shapiro_p))

@@

```{r results="asis", echo=FALSE}
## Print table
kbl(results, format = "latex", booktabs = T, caption = "Summary of SE estimates for mediation effect under continuous response, continuous mediator, fixed-effects model \\label{tab:results_bin_bin_mix-Shapiro}")
```
